---
title: "Replicate Concordance"
author: "JL"
date: "7/12/2017"
output: 
  html_document: 
    keep_md: yes
---

TODO 

- manual
- automatic
  - with/without QC
  
1. CV of reps
2. ICC
3. min diff
4. max diff
  
```{r setup,eval=FALSE, include=FALSE}
mets =   read.delim("/Volumes/Beta/data/flow/concordance/results_r6/freq.metrics.txt",stringsAsFactors = FALSE,sep = "\t") 
save(mets,file ="/Volumes/Beta/data/flow/concordance/results_r6/freq.metrics.Rdata" )

```

```{r eval=TRUE, echo=FALSE,warning=FALSE,include=FALSE}
source(file = "/Users/Kitty/git/auto-fcs/explore/openCyto/computeFreqs.R")
library(knitr)
library(raster)
library("ggplot2")
library(plyr)
theme_set(theme_bw(20))

load("/Volumes/Beta/data/flow/concordance/results_r6/freq.metrics.Rdata")
```

```{r eval=TRUE, echo=FALSE}
map = read.delim("/Users/Kitty/git/auto-fcs/explore/openCyto/panel1Map.txt",stringsAsFactors = FALSE,sep = "\t") 
metsLymph = mets[which(mets$Population=="lymph"),]
mets50K = metsLymph[which(metsLymph$Count>50000),]
ctrlUse =grepl("_Ctl",mets$name)
metsControls =mets[ctrlUse,]
metsControls$CTRL_NAME = gsub(".*_Ctl","",metsControls$name)
metsControls$CTRL_NAME = gsub("^_","",metsControls$CTRL_NAME)

metsControls$CTRL_NAME = gsub("_.*","",metsControls$CTRL_NAME)
# kable(as.data.frame(table(metsControls$CTRL_NAME)),format = "markdown")
metsControls$CTRL_NAME = gsub("-","",metsControls$CTRL_NAME)
metsControls$CTRL_NAME = gsub("DHS","",metsControls$CTRL_NAME)
metsControls$CTRL_NAME = gsub("newbiolegend","",metsControls$CTRL_NAME)
metsControls$CTRL_NAME =paste(metsControls$CTRL_NAME ,metsControls$Panel,"QC",metsControls$QC,sep="_" )
metsControls$Population =gsub("FITC-A+BV711-A-","BB515-A+BV 711-A-",metsControls$Population,fixed = TRUE)
metsControls$Population =gsub("FITC-A-BV711-A-","BB515-A-BV 711-A-",metsControls$Population,fixed = TRUE)
metsControls$Freq_Base=NA
metsControls$Freq_Base[metsControls$Panel=="panel1"] =metsControls$freq_lymph[metsControls$Panel=="panel1"]
metsControls$Freq_Base[metsControls$Panel=="panel2"] =metsControls$freq_PBMC[metsControls$Panel=="panel2"]

metsControls$Population  =paste(metsControls$Population ,metsControls$Panel ,sep="_" )
# metsControlsAuto =computFreqs(metsD =metsControls,panel = "panel1",map = map,basePop = "lymph",QC = "FALSE" )
# 
metsControlsAuto =metsControls
# kable(as.data.frame(table(metsControlsAuto$CTRL_NAME)),format = "markdown")
sum =as.data.frame(table(metsControlsAuto[which(metsControlsAuto$Parent=="root"),]$CTRL_NAME))
colnames(sum) =c("REPLICATE","NUMBER_FCS_FILES")
kable(sum,format = "markdown")

```


Prop of  files with lymph counts > 50K  = `r length(mets50K$name)/length(metsLymph$name)`
Prop of  files with lymph counts < 50K  = `r 1-(length(mets50K$name)/length(metsLymph$name))`

Total number of Control fcs files (currently) = `r length(unique(metsControls$name))`

Total number of Control groups (currently) = `r length(unique(metsControls$CTRL_NAME))`

```{r eval=TRUE, echo=FALSE,fig.width=10,fig.height=8,warning=FALSE}


df = data.frame()

for(pop in unique(metsControlsAuto$Population)) {
  for (group in unique(metsControlsAuto$CTRL_NAME)) {
    sub = metsControlsAuto[which(metsControlsAuto$Population == pop &
                              metsControlsAuto$CTRL_NAME == group),]
    # print(paste(pop,group,))
    #     print(paste(pop,group,cv(sub$Count,na.rm = TRUE)))
    tmp = data.frame(
      POP = pop,
      GROUP = group,
      CV_FREQ_PARENT = cv(sub$freqParent,na.rm = TRUE),
      CV_FREQ_BASE = cv(sub$Freq_Base,na.rm = TRUE),
      QC = grepl(pattern = "TRUE",x = group)
      # ,
      # PANEL = unique(sub$Panel)
    )
    df = rbind.data.frame(df, tmp)

  }
}
p <- ggplot(df, aes(x=POP, y=CV_FREQ_PARENT,color=QC))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("CV Group (freq parent)")+xlab("population") +facet_grid(QC ~ .)

p <- ggplot(df, aes(x=POP, y=CV_FREQ_BASE,color=QC))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("CV Group (freq lymph/pbmc)")+xlab("population")+facet_grid(QC ~ .)

```

Percent of parent CV summary

```{r eval=TRUE, echo=FALSE,fig.width=10,fig.height=8,warning=FALSE}
df$GROUP_MEDIAN_CV = df$CV_FREQ_PARENT
df$GROUP_MAD_CV = df$CV_FREQ_PARENT

meds= aggregate(GROUP_MEDIAN_CV ~ GROUP, df, FUN =median,na.action = na.omit)
mads= aggregate(GROUP_MAD_CV ~ GROUP, df, FUN =mad,na.action = na.omit)

p <- ggplot(df, aes(x=POP, y=CV_FREQ_PARENT,color=QC))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("CV Group (freq parent)")+xlab("population") +facet_grid(QC ~ .)

kable(merge(meds,mads,by.x = "GROUP",by.y = "GROUP"),format = "markdown")

```


Percent of base population (lymph/PBMC) CV summary

```{r eval=TRUE, echo=FALSE,fig.width=10,fig.height=8,warning=FALSE}
df$GROUP_MEDIAN_CV = df$CV_FREQ_BASE
df$GROUP_MAD_CV = df$CV_FREQ_BASE

meds= aggregate(GROUP_MEDIAN_CV ~ GROUP, df, FUN =median,na.action = na.omit)
mads= aggregate(GROUP_MAD_CV ~ GROUP, df, FUN =mad,na.action = na.omit)

kable(merge(meds,mads,by.x = "GROUP",by.y = "GROUP"),format = "markdown")

for (group in unique(metsControlsAuto$CTRL_NAME)) {
  sub = metsControlsAuto[which(
  metsControlsAuto$CTRL_NAME == group), ]
  p <- ggplot(sub, aes(x = Population, y = 100*freqParent))+ geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust =
  1)) + ylab("% of parent") + xlab("population") + ggtitle(paste("Ctl",group,"n =",length(unique(sub$name))))
  print(p)
  p <- ggplot(sub, aes(x = Population, y = 100*Freq_Base))+ geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust =
  1)) + ylab("% of base pop (lymph/pbmc)") + xlab("population") + ggtitle(paste("Ctl",group,"n =",length(unique(sub$name))))
  print(p)
}


```

# Replicate data clustering


```{r eval=TRUE, echo=FALSE,fig.width=10,fig.height=8,warning=FALSE}

library(reshape2)
library(superheat)

metsControlsFilt =metsControlsAuto[which(metsControlsAuto$Population!="boundary"),]
freqpart = dcast(metsControlsFilt,name~Population,value.var ="freqParent" )
row.names(freqpart) = freqpart$name
freqpart =freqpart[,c(-1)]
# 
# freq.pca <- prcomp(freqpart,center = TRUE,scale = TRUE) 
# pcaDat = as.data.frame(freq.pca$x)
# pcaDat$CTRL_NAME = gsub(".*_Ctl","",row.names(pcaDat))
# pcaDat$CTRL_NAME = gsub("_G","G",pcaDat$CTRL_NAME)
# pcaDat$CTRL_NAME = gsub("_.*","",pcaDat$CTRL_NAME)
# # kable(as.data.frame(table(metsControls$CTRL_NAME)),format = "markdown")
# pcaDat$CTRL_NAME = gsub("-","",pcaDat$CTRL_NAME)
# pcaDat$CTRL_NAME = gsub("DHS","",pcaDat$CTRL_NAME)
# pcaDat$CTRL_NAME = gsub("newbiolegend","",pcaDat$CTRL_NAME)
# 
# 
# p = ggplot(pcaDat,aes(x=PC1,y=PC2, color=CTRL_NAME)) +geom_point()+xlab("PC1")+ylab("PC2")
# p 
# 
# p = ggplot(pcaDat,aes(x=PC3,y=PC2, color=CTRL_NAME)) +geom_point()+xlab("PC3")+ylab("PC2")
# p 



# 
# clustDat = scale(as.matrix(freqpart),center = TRUE,scale = TRUE)
# superheat(freqpart, 
# 
#           # place dendrograms on columns and rows 
#           row.dendrogram = T, 
#           col.dendrogram = T,
#           
#           # make gridlines white for enhanced prettiness
#           grid.hline.col = "white",
#           grid.vline.col = "white",
#           
#           # rotate bottom label text
#           bottom.label.text.angle = 90)
# 
# 
# hlustfunc <- function(x) hclust(x, method="complete")
# distfunc <- function(x) as.dist((1-cor(t(x)))/2)
# d <- distfunc(clustDat)
# fit <- hlustfunc(d)
# # plot(fit)
# 
# clusterCut <- as.data.frame(cutree(fit, 8))
# clusterCut= clusterCut[order(clusterCut$`cutree(fit, 8)`),] 


# kable(clusterCut,format = "markdown")

```