---
title: "results_r18 LOD"
author: "JL"
date: "10/10/2017"
output: 
  html_document: 
    keep_md: yes
---

##  LOD testing



```{r setup,echo=FALSE}
library(ggplot2)
library(knitr)
theme_set(theme_bw(20))
t2 <- theme(
    axis.line = element_line(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(angle=90,hjust=1)
)

gates =   read.delim(
  "/Volumes/Beta/data/flow/results_r18/LOD/all.freq.metrics.txt",
  stringsAsFactors = FALSE,
  sep = "\t"
  ) 
gates$key=paste(gates$name,gates$Panel)

counts =   read.delim(
  "/Volumes/Beta/data/flow/results_r18/LOD/all.totalCellCounts.metrics.txt",
  stringsAsFactors = FALSE,
  sep = "\t"
  ) 
counts$PLOT_NAME=counts$FILE

counts$PLOT_NAME=gsub("2017-10-05_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("LSR_ZF_LOD_Group one_Ctl-G_LOD_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("2017-03-02_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("ZF_group two_Ctl-G_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("DHS_group two_Ctl-G_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("FORTESSA_HB_group two_Ctl-G_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("2017-03-07_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("2017-02-13_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("2017-02-27_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("_group two_Ctl_G_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("_Group two_Ctl-G_","",counts$PLOT_NAME)
counts$PLOT_NAME=gsub("PANEL ","",counts$PLOT_NAME)


counts$key=paste(counts$FILE,counts$PANEL)

# gates = merge(gates,counts,by.x = "key",by.y = "key",all.y = TRUE)

counts$PANEL=as.factor(counts$PANEL)


```

- total of `r length(counts$PANEL)` fcs files

```{r echo=FALSE,fig.height=10,fig.width=12}

counts$SUCCESS =counts$FILE %in% gates$name
ggplot(data=counts, aes(x=PLOT_NAME, y=TOTAL_COUNTS, fill=counts$PANEL)) +
    geom_bar(colour="black", stat="identity") +t2+xlab("Trimmed fcs name")

```

- TOTAL_COUNTS is the total number of cells in the .fcs file
- one sample had an "undetermined panel" , meaning the .fcs file was probably missing a desc column


```{r echo=FALSE,fig.height=10,fig.width=12}

counts$SUCCESS_PANEL =as.factor(paste0(counts$PANEL,counts$SUCCESS))

ggplot(data=counts, aes(x=PLOT_NAME, y=TOTAL_COUNTS, fill=counts$SUCCESS_PANEL)) +
    geom_bar(colour="black", stat="identity") +t2+xlab("Trimmed fcs name")
```

- panel1TRUE means the sample was successfully gated
- panel1FALSE means the sample was not successfully gated
- panel2TRUE means the sample was successfully gated
- panel2FALSE means the sample was not successfully gated


```{r echo=FALSE,fig.height=10,fig.width=12}
fails =counts[which(counts$SUCCESS==FALSE),]
ggplot(data=fails, aes(x=PLOT_NAME, y=TOTAL_COUNTS, fill=fails$SUCCESS_PANEL)) +
    geom_bar(colour="black", stat="identity") +t2+xlab("Trimmed fcs name")
```



- only the samples that failed gating

### File name mappings 

```{r echo=FALSE,fig.height=10,fig.width=12}

kable(counts[,c("FILE","PLOT_NAME","TOTAL_COUNTS")],format = "markdown")

```


