---
title: "OpenCyto"
author: "JL"
date: "7/19/2017"
output: 
  html_document: 
    keep_md: yes
---

Almost Documentation ? 

```{r pressure, echo=FALSE}
library(readr)
batch <- read_file("~/git/auto-fcs/explore/openCyto/msiScripts/runGateMSI")
assemble <- read_file("~/git/auto-fcs/explore/openCyto/msiScripts/collectResultsMSI")


```

## Running OpenCyto

1. Set up batches with genvisis

```bash
`r batch`
```
  - Result directories will be created in `$OUTDIR` that correspond to each batch
    - e.g `$OUTDIR/openCytoBatch_0/` ,`$OUTDIR/openCytoBatch_1/`, etc
2. Submit all .pbs scripts in `$OUTDIR`
    - e.g `qp *.pbs`
    - Current runtime on for 16 batches on 2 pankrat q nodes is ~7hrs without QC, 37hrs with QC
    
## OpenCyto results

Each batch output directory (as an example, `$OUTDIR/openCytoBatch_0/`) should contain the following:

- `gates/`: **directory containing .wsp files for each successfully gated sample**
    - each sample will have 2 .wsp files, `sampleName_panelX.wsp`, and `sampleName_panelXRename.wsp`
        - `sampleName_panelXRename.wsp` contains gates named according to the manual hierarchy
- `gatesQC/`: **directory containing .wsp files for each successfully gated sample, after QC processing**
     - also will have 2 .wsp files per sample, `sampleName_panelX.wsp`, and `sampleName_panelXRename.wsp`

- `fcsQC/`: **directory containing .fcs files representing HQ events, currently according to FlowAI**
     - in order to reproduce the counts listed in gatesQC .wsp files, these files should be used

- `gate_plotsXX.pdf`: **Plots of each OpenCyto gate**

```{r headers, echo=FALSE}
library(knitr)
total = data.frame(HEADER = c("FILE","TOTAL_COUNTS","QC","PANEL"))

```

- `metrics.totalCellCounts.txt`: **File containing total cell counts for all samples**
    - **FILE** : fcs file
    - **TOTAL_COUNTS**: total number of events in the fcs file
    - **QC** : whether QC was performed prior to counting
    - **PANEL** :  panel analyzed
    
- `metrics.txt`: **Gate metrics for each succesfully gated sample**
    - **name**         : fcs file
    - **Population**   : Cell population
    - **Parent**       : Parent population
    - **Count**        : Event count for  population
    - **ParentCount**  : Event count for parent population
    - **METRIC**       : should be always counts
    - **GATE**         : Whether we are currently reporting automatic (openCyto) or manual gating results
    - **QC**           : Whether QC was applied prior to gating
    - **Panel**           : panel1 or panel2
    - **PDF**           : name of the pdf containing plots for this sample
    - **FlaggedSample**           : placeholder to flag problematic samples for review

- `freq.metrics.txt`: **Similar to `metrics.txt`, but with additional info**
     - **RealAutoParent** : when the `Parent` gate is actually an intermediate step, this lists the actual parent population
      - **Manual**   : Gate name in manual nomenclature
      - **Merge**   : Data key of **population_sample**
      - **freq_lymph**   : for panel1 samples, the frequency of the population relative to lymphocytes, and `NA` for panel2 samples
      - **freqParent**   : for panel1 and panel2 samples, the frequency of the population relative to the `RealAutoParent` population
      - **freq_PBMC**   : for panel2 samples, the frequency of the population relative to PBMCs, and `NA` for panel1 samples
      
      
To assemble results files from each batch, run the following from within `$OUTDIR`:

```bash
`r assemble`
```
