---
title: "Review OpenCyto fails"
author: "JL"
date: "6/30/2017"
output: 
  html_document: 
    keep_md: yes
---
### NOTES:
 - FORTESSA determined by filename, nothing else
 - date determined by filename, nothing else


```{r setup,eval=FALSE, include=FALSE}
counts =  read.delim("/Volumes/Beta/data/flow/checkFails/all.totalCellCounts.metrics.txt",stringsAsFactors = FALSE,sep = "\t") 
mets =   read.delim("/Volumes/Beta/data/flow/checkFails/all.metrics.txt",stringsAsFactors = FALSE,sep = "\t") 
save(mets,file ="/Volumes/Beta/data/flow/checkFails/all.metrics.Rdata" )
save(counts,file ="/Volumes/Beta/data/flow/checkFails/all.totalCellCounts.metrics.Rdata" )

```

### SUCCESSFUL_GATING Cell counts: All samples

```{r echo=FALSE,fig.width=10}
# ,fig.height=8
library(knitr)

library(ggplot2)
theme_set(theme_bw(20))
skip =  read.delim("/Volumes/Beta/data/flow/checkFails/axe.txt",stringsAsFactors = FALSE,sep = "\t",header = FALSE) 
qcMetrics =  read.delim("/Volumes/Beta/data/flow/QCeval/all.mini.qc.txt",stringsAsFactors = FALSE,sep = "\t") 
qcMetrics$Name.file =paste(qcMetrics$Name.file,".fcs",sep = "")
load("/Volumes/Beta/data/flow/checkFails/all.metrics.Rdata")
load("/Volumes/Beta/data/flow/checkFails/all.totalCellCounts.metrics.Rdata")
# counts = counts[which(counts$PANEL=="panel1"&counts$QC=="FALSE"),]
counts = merge(counts,qcMetrics,by.x = "FILE",by.y ="Name.file",all.x = TRUE )
NO_DESC=counts$FILE %in% skip$V1
complete = counts$FILE %in% mets$name
counts$SUCCESSFUL_GATING =complete 
counts$NO_DESC =NO_DESC
counts$NO_DESC =gsub("TRUE","NO_DESC",counts$NO_DESC )
counts$NO_DESC =gsub("FALSE","DESC",counts$NO_DESC )

counts$LOG_TOTAL_COUNT = log10(counts$TOTAL_COUNTS)

counts$PROCESS_DATE = gsub("PANEL 1.*","",counts$FILE)
counts$FORTESSA = gsub(".*FORTESSA.*","FORTESSA",counts$FILE,fixed = FALSE)
counts$FORTESSA = gsub(".*PANEL 1.*","NOT_FORTESSA",counts$FORTESSA,fixed = FALSE)


counts$PROCESS_DATE =paste(counts$PROCESS_DATE,counts$FORTESSA)

counts = counts[which(counts$FILE!="2016-05-03_P2_HB_panel 2_CTL-A_1715_001.fcs"&counts$FILE!="Specimen_001_WHOLE BLOOD BT RBC X1_014.fcs"&counts$FILE!="Specimen_001_WHOLE BLOOD BT RBC X2_015.fcs"&counts$FILE!="2016-05-03_P2_HB_panel 2_CTL-B_1722_002.fcs"),]

ggplot(counts, aes(LOG_TOTAL_COUNT,colour = SUCCESSFUL_GATING)) +
  geom_histogram(bins = 300)+ylab("Number of Samples")+xlab("Log10 total cell count") +facet_wrap(~SUCCESSFUL_GATING)



```

### SUCCESSFUL_GATING Cell counts: Split by machine


```{r echo=FALSE,fig.width=10}
ggplot(counts, aes(LOG_TOTAL_COUNT,colour = SUCCESSFUL_GATING)) +
  geom_histogram(bins = 300)+ylab("Number of Samples")+xlab("Log10 total cell count") +facet_wrap(FORTESSA~SUCCESSFUL_GATING)


fails =counts[which(counts$SUCCESSFUL_GATING==FALSE),]
pfail =length(fails$FILE)/length(counts$FILE)

frT = 5
t = as.data.frame(table(counts$PROCESS_DATE,counts$SUCCESSFUL_GATING))
tf = t[which(t$Var2==FALSE),]
tt = t[which(t$Var2==TRUE),]

tm = merge(tf,tt,by.x = "Var1",by.y = "Var1")
tm$PERCENT_FAIL_BY_BATCH =100* tm$Freq.x/(tm$Freq.y+tm$Freq.x)
tm$FAIL_RANK = rank(tm$PERCENT_FAIL_BY_BATCH)

counts =  merge(counts,tm,by.x = "PROCESS_DATE",by.y = "Var1")
counts$FAIL_RANK = rank(counts$PERCENT_FAIL_BY_BATCH)
countsGT10 = counts[which(counts$PERCENT_FAIL_BY_BATCH>frT),]
countsLT10 = counts[which(counts$PERCENT_FAIL_BY_BATCH<=frT),]

```

### SUCCESSFUL_GATING Cell counts: Split by files with no desc column


```{r echo=FALSE,fig.width=10}
ggplot(counts, aes(LOG_TOTAL_COUNT,colour = SUCCESSFUL_GATING)) +
  geom_histogram(bins = 300)+ylab("Number of Samples")+xlab("Log10 total cell count") +facet_wrap(NO_DESC~SUCCESSFUL_GATING)

kable(as.data.frame(table(counts$SUCCESSFUL_GATING,counts$NO_DESC)),format = "markdown")
```



### SUCCESSFUL_GATING Cell counts: Only Batches with >`r frT`% failure rate (`r length(countsGT10$FILE)` total samples)


```{r echo=FALSE,fig.width=10}

# 
ggplot(countsGT10, aes(LOG_TOTAL_COUNT,colour = SUCCESSFUL_GATING)) +
  geom_histogram(bins = 300)+ylab("Number of Samples")+xlab("Log10 total cell count") +facet_wrap(~SUCCESSFUL_GATING)

```

### SUCCESSFUL_GATING Cell counts: Only Batches with <=`r frT`% failure rate (`r length(countsLT10$FILE)` total samples)


```{r echo=FALSE,fig.width=10}
ggplot(countsLT10, aes(LOG_TOTAL_COUNT,colour = SUCCESSFUL_GATING)) +
  geom_histogram(bins = 300)+ylab("Number of Samples (truncated)")+xlab("Log10 total cell count") +facet_wrap(~SUCCESSFUL_GATING) +ylim(0,50)
ggplot(countsLT10, aes(LOG_TOTAL_COUNT,colour = SUCCESSFUL_GATING)) +
  geom_histogram(bins = 300)+ylab("Number of Samples")+xlab("Log10 total cell count") +facet_wrap(~SUCCESSFUL_GATING)
```

### SUCCESSFUL_GATING QC comparison


 All event counts
```{r echo=FALSE,fig.width=10}
p <- ggplot(counts, aes(x=SUCCESSFUL_GATING, y=X..anomalies.flow.Rate,color=SUCCESSFUL_GATING))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("percent anomalies.flow.Rate")+xlab("success")
p <- ggplot(counts, aes(x=SUCCESSFUL_GATING, y=X..anomalies.Signal,color=SUCCESSFUL_GATING))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("percent anomalies.Signal")+xlab("success")
p <- ggplot(counts, aes(x=SUCCESSFUL_GATING, y=X..anomalies.Margins,color=SUCCESSFUL_GATING))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("percent anomalies.Margins")+xlab("success")



```



Event counts >100k
```{r echo=FALSE,fig.width=10}
p <- ggplot(counts[which(counts$TOTAL_COUNTS>100000),], aes(x=SUCCESSFUL_GATING, y=X..anomalies.flow.Rate,color=SUCCESSFUL_GATING))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("percent anomalies.flow.Rate")+xlab("success")
p <- ggplot(counts[which(counts$TOTAL_COUNTS>100000),], aes(x=SUCCESSFUL_GATING, y=X..anomalies.Signal,color=SUCCESSFUL_GATING))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("percent anomalies.Signal")+xlab("success")
p <- ggplot(counts[which(counts$TOTAL_COUNTS>100000),], aes(x=SUCCESSFUL_GATING, y=X..anomalies.Margins,color=SUCCESSFUL_GATING))
p + geom_boxplot() +theme(axis.text.x=element_text(angle=90,hjust=1)) +ylab("percent anomalies.Margins")+xlab("success")

```



### SUCCESSFUL_GATING NOTES

***`r pfail*100`% failure rate**



```{r echo=FALSE,fig.width=10}

kable(tm[order(-tm$PERCENT_FAIL_BY_BATCH),],format = "markdown")

table(counts$SUCCESSFUL_GATING)

kable(fails[order(-fails$TOTAL_COUNTS),],format = "markdown")

tmFail = fails[which(fails$PROCESS_DATE=="2017-01-09_ NOT_FORTESSA"),]

for(file in tmFail$FILE){
  dir ="/Volumes/Beta/data/flow/checkFails/badBatch/"
  qcFile = paste(dir,gsub(".fcs","",file,fixed = TRUE),"_QC.html",sep="")
  qcFailFile = paste(dir,"fails/",gsub(".fcs","",file,fixed = TRUE),"_QC.html",sep="")

  file.copy(qcFile,qcFailFile)
  
}

```