---
title: "FlowAI"
author: "JL"
output: html_document
---

# 

```{r setusp, include=FALSE}
library(openCyto)
library(flowCore)
library(data.table)
library(ggcyto)
library(gridExtra)
library(CytoML)
library(flowAI)

# TP(Color.GREEN),
# 		TN(Color.BLACK),
# 		FP(Color.BLUE),
# 		FN(Color.RED);

```

```{r warning=FALSE,results='hide',echo=FALSE}
inputFCSDir = "/Volumes/Beta/data/flow/fcs3/"
cleanDir = "/Volumes/Beta/data/flow/fcs3Clean/"
setwd(cleanDir)
outputDir = "/Volumes/Beta/data/flow/"
template = "~/git/auto-fcs/explore/openCyto/lymph.dev.b.csv"
templateQC = "~/git/auto-fcs/explore/openCyto/lymph.dev.qc.csv"

gateDir = "gates5/"
gateQCDir = "gatesQC5/"

# getIndiceMat()
theme_set(theme_bw(5))
gt_lymph <-
  gatingTemplate(template, autostart = 1L)
gt_lymphQC <-
  gatingTemplate(templateQC, autostart = 1L)
fcsFilesAll <-
  list.files(inputFCSDir,
             pattern = ".fcs",
             full = FALSE)
fcsFilesAll =sample(fcsFilesAll,20)
fcsFilesAllProbs = c(
  "2016-05-29_PANEL 1_HB_panel one_H_D3_003.fcs", #Panel 1 test 6, strange clustering
  "2016-08-01_PANEL 1_DHS_Group one_F1630927_008.fcs",
  "2016-05-06_PANEL 1_ZF_panel one_F1631950_006.fcs", #test 1 page 7 
  "2016-08-01_PANEL 1_DHS_Group one_F1636819_011.fcs",
  "2016-08-01_PANEL 1_DHS_Group two_F1631374_028.fcs",
  "2016-05-27_PANEL 1_ZF_panel one_F1631054_011.fcs", #test 5 page 4  weird additional population of cells 
  "2016-05-27_PANEL 1_ZF_panel one_F1631061_005.fcs" ,#test 5 page 6  weird additional population of cells
  "2016-05-27_PANEL 1_ZF_panel one_F1631137_009.fcs",#test 5 page 8 lower cluster below the lymphocytes
  "2016-05-27_PANEL 1_ZF_panel one_F1631355_008.fcs" ,#test 5 page 8 lower cluster below the lymphocytes
  "2016-08-01_PANEL 1_DHS_Group two_F1631374_028.fcs",# most over -called
  "2016-08-01_PANEL 1_DHS_Group one_F1636819_011.fcs",# 2nd most over -called
  "2016-07-29_PANEL 1_ZF_Group one_F1632280_001.fcs", # most under -called
  "2016-07-29_PANEL 1_ZF_Group two_F1636833_030.fcs", # 2nd most under -called
  "2016-07-08_PANEL 1_ZF_Group one_F1632339_009.fcs" #Crazy flowrate
  )

fcsFilesAll =unique(c(fcsFilesAllProbs,fcsFilesAll))
fcsFilesAll = split(fcsFilesAll, ceiling(seq_along(fcsFilesAll) / 5))
# 19:49 -> 20:31
# [30:33]


compFrame <- function(frame, file, gt_lymph, d, outputDir, gateDir,qcVersion) {
      print(paste("compensating ....", file))

      comp <- compensation(keyword(frame)$`SPILL`)
      
      chnls <- parameters(comp)
      biexpTrans <-
        flowJo_biexp_trans(
          channelRange = 256,
          maxValue = 262144.0000291775 ,
          pos = 4.418539922,
          neg = 0,
          widthBasis = -100
        )
      tf <- transformerList(chnls, biexpTrans)
      
      print(paste("gating ....", file))
      # gs <- GatingSet(c(new_frame))
      frames = c(frame)
      names(frames) = c(basename(file))
      fs =  as(frames, "flowSet")
      gs1 <- GatingSet(fs)
      gs1 <- compensate(gs1, comp)
      gs1 <- transform(gs1, tf)
      
      gh <- gs1[[1]]
      gating(gt_lymph, gs1)
      print(paste("plotting ....", file))
      
      t1 = ggcyto(gs1,
                  mapping = aes(x = "FSC-A", y = "SSC-A"),
                  subset = "root") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate() + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
      
      # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
      
      t2 = ggcyto(gs1,
                  mapping = aes(x = "FSC-A", y = "SSC-A"),
                  subset = "boundary") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate() + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
      # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
      
      # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
      if (qcVersion) {
        t2_1 = ggcyto(gs1,
        mapping = aes(x = "FSC-A", y = "SSC-A"),
        subset = "QC") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate() + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
        # + scale_x_flowJo_biexp() + scale_y_flowJo_biexp()
      }else{
        t2_1 =t2
      }
       # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
       
      t3 = ggcyto(gs1,
                  mapping = aes(x = "FSC-A", y = "SSC-A"),
                  subset = "nonDebris") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate() + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
      # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
      
      
      t4 = ggcyto(gs1,
                  mapping = aes(x = "FSC-W", y = "FSC-H"),
                  subset = "lymph") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
      
      t5 = ggcyto(gs1,
                  mapping = aes(x = "PE-A", y = "FSC-H"),
                  subset = "Singlets") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate() 
      
      t6 = ggcyto(gs1,
                  mapping = aes(x = "CD3", y = "CD19"),
                  subset = "PE-A") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
      
      t1Bcell = ggcyto(gs1,
                  mapping = aes(x = "CD3", y = "CD19"),
                  subset = "CD3-") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
      
      
      
       t7 = ggcyto(gs1,
                  mapping = aes(x = "CD4", y = "CD8"),
                  subset = "CD3+") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
      
        t7_1 = ggcyto(gs1,
                  mapping = aes(x = "CD4", y = "CD8"),
                  subset = "CD4_1") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
      
       t8 = ggcyto(gs1,
                  mapping = aes(x = "CD4", y = "HLA-DR"),
                  subset = "CD4") +
        geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
       
       t9 = ggcyto(gs1,mapping = aes(x = "CD45RA",y = "CCR7"),subset = "CD4") +geom_hex(bins=100) + ggcyto_par_set(limits = "data") + geom_gate()
       
      grid.arrange(
        as.ggplot(t1),
        as.ggplot(t2),
        
        as.ggplot(t2_1),
  
        as.ggplot(t3),
        as.ggplot(t4),
        as.ggplot(t5),
        as.ggplot(t6),
        as.ggplot(t1Bcell),
        as.ggplot(t7),
        as.ggplot(t7_1),
        as.ggplot(t8),
        as.ggplot(t9),

        ncol = 2
      )
   
      
      d = rbind.data.frame(as.data.frame(getPopStats(gs1)), d)
      outFile <-
        paste(outputDir,gateDir,
              basename(file),
              "_panel1.wsp", sep = "")
      GatingSet2flowJo(gs1, outFile)
      
  
      gateDef = data.frame(
        lymph = getIndiceMat(gs1, "lymph"),
        Singlets = getIndiceMat(gs1, "Singlets"),
        PE_A_Minus = getIndiceMat(gs1, "PE-A"),
        CD3Plus = getIndiceMat(gs1, "CD3+"),
        cd4Pcd8M = getIndiceMat(gs1, "CD4"),
        cd4Mcd8P = getIndiceMat(gs1, "CD8"),
        HLADR = getIndiceMat(gs1, "ActivatedCD4"),
        CCR7PCD45RAP = getIndiceMat(gs1, "CCR7+CD45RA+"),
        CCR7PCD45RAM = getIndiceMat(gs1, "CCR7+CD45RA-"),
        CCR7MCD45RAM = getIndiceMat(gs1, "CCR7-CD45RA-"),
        CCR7MCD45RAP = getIndiceMat(gs1, "CCR7-CD45RA+"),
        BCELL = getIndiceMat(gs1, "CD19+")

        )
      colnames(gateDef) = c(
        "lymph",
        "Singlets",
        "PE_A",
        "CD3_PLUS",
        "CD4",
        "CD8",
        "ActivatedCD4",
        "CCR7P_CD45RAP",
        "CCR7P_CD45RAM",
        "CCR7M_CD45RAM",
        "CCR7M_CD45RAP",
        "BCELL"
        )
      # APC-Cy7-A = CD4
      # BUV 395-A =CD8
      
      write.table(
        gateDef,
        sep = "\t",
        quote = FALSE,
        file = paste(
          outputDir,
          gateDir,
          basename(file),
          "_panel1_gate_def.txt",
          sep = ""
        ),
        row.names = FALSE
      )

    }



# https://www.bioconductor.org/packages/devel/bioc/vignettes/flowCore/inst/doc/HowTo-flowCore.pdf
i = 1
numProcessed =0
d = data.frame()
for (files in fcsFilesAll) {
  # fcsFiles = files
  print(files)
  i
  pdf(file = paste(outputDir, "panel1_test5_", i, ".pdf", sep = ""))
  i = i + 1
  for (file in files) {
    numProcessed = numProcessed+1
    print(paste("number processed ....", numProcessed))
    print(paste("loading ....", file))
    frame = read.FCS(paste(inputFCSDir,file,sep = ""))
    description(frame)$FILENAME =file
    
    qcFile = paste(tools::file_path_sans_ext(file), ".fcs", sep = "")
    qcFileFull =paste(cleanDir,"QC/",qcFile,sep = "")
    if(!file.exists(qcFileFull)){
    flow_auto_qc(
      frame,
      folder_results = "QC",
      mini_report = paste(basename(file), "mini"),
      fcs_QC = "",
      pen_valueFS = 50,
      remove_from = "FR_FM"
      )
    }
    compFrame(
      frame = frame,
      file = file,
      gt_lymph = gt_lymph ,
      d = d,
      outputDir = outputDir,
      gateDir = gateDir,
      qcVersion = FALSE
      )

    # frame.c = read.FCS(qcFileFull)
    # compFrame(
    #   frame = frame.c,
    #   file = file,
    #   gt_lymph = gt_lymphQC ,
    #   d = d,
    #   outputDir = outputDir,
    #   gateDir = gateQCDir,
    #   qcVersion = TRUE
    #   )
  }
  
  dev.off()
}




write.table(
  d,
  sep = "\t",
  quote = FALSE,
  file = paste(outputDir, "testTypesTest5.txt", sep = ""),
  row.names = FALSE
)
# https://github.com/RGLab/openCyto/issues/110autoplot(gs[[1]])

```


ws <- openWorkspace("/Users/Kitty/temp/fcs/513-520 HRS p1&p2.wsp")
 gs <- parseWorkspace(ws, path = "/Volumes/Beta/data/flow/fcs/", name = 1,subset="2016-05-18_PANEL 1_ZF_panel one_F1632164_009.fcs",isNcdf = FALSE)
t3 = ggcyto(gs,
            mapping = aes(x = "FSC-A", y = "SSC-A"),
            subset = "Lymphocytes (SSC-A v FSC-A)") +
    geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate() + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
write.table(
+     getPopStats(gs),
+     sep = "\t",
+     quote = FALSE,
+     file = "/Volumes/Beta/data/flow/testTypesManual.txt",
+     row.names = FALSE
+ )

```

   
      gt_lymph = rbind(
        gt_lymph,
        add_pop(
        gs1,
        alias = "*",
        pop = "HLADR+/-",
        parent = "CD4",
        dims =  "HLA-DR",
        gating_method = "tailgate",
        gating_args =
        "tol=5e-2",
        preprocessing_method = "standardize_flowset"
        )
        )
      
      gt_lymph = rbind(
        gt_lymph,
        add_pop(
        gs1,
        alias = "*",
        pop = "CD45RA+/-",
        parent = "CD4",
        dims =  "CD45RA",
        gating_method = "mindensity",
        gating_args =
        "gate_range=c(80,150)"
        )
        )
      gt_lymph = rbind(
  gt_lymph,
  add_pop(
    gs1,
    alias = "CCR7_gate",
    pop = "CCR7",
    parent = "CD45RA+",
    dims =  "CCR7",
    gating_method = "flowClust",
    gating_args = "K=2",
    preprocessing_method = "prior_flowClust",
    preprocessing_args = "K=2,min=0"
  )
)

