---
title: "FlowAI"
author: "JL"
output: html_document
---

# 

```{r setusp, include=FALSE}
library(openCyto)
library(flowCore)
library(data.table)
library(ggcyto)
library(gridExtra)
library(CytoML)
library(flowAI)

# TP(Color.GREEN),
# 		TN(Color.BLACK),
# 		FP(Color.BLUE),
# 		FN(Color.RED);

```

```{r warning=FALSE,results='hide',echo=FALSE}
inputFCSDir = "/Volumes/Beta/data/flow/fcs3/"
cleanDir = "/Volumes/Beta/data/flow/fcs3Clean/"
setwd(cleanDir)
outputDir = "/Volumes/Beta/data/flow/"
template = "~/git/auto-fcs/explore/openCyto/lymph.dev.b.csv"
templateQC = "~/git/auto-fcs/explore/openCyto/lymph.dev.qc.b.csv"

gateDir = "gates5/"
gateQCDir = "gatesQC5/"
mapper =read.delim("/Volumes/Beta/data/flow/fcsMap.txt",stringsAsFactors = FALSE,sep = "\t") 

# getIndiceMat()
theme_set(theme_bw(5))
gt_lymph <-
  gatingTemplate(template, autostart = 1L)
gt_lymphQC <-
  gatingTemplate(templateQC, autostart = 1L)
fcsFilesAll <-
  list.files(inputFCSDir,
             pattern = ".fcs",
             full = FALSE)
fcsFilesAll = sample(fcsFilesAll, 10)
fcsFilesAllProbs = c(
  "2016-05-29_PANEL 1_HB_panel one_H_D3_003.fcs", #Panel 1 test 6, strange clustering
  "2016-08-01_PANEL 1_DHS_Group one_F1630927_008.fcs",
  "2016-05-06_PANEL 1_ZF_panel one_F1631950_006.fcs", #test 1 page 7 
  "2016-08-01_PANEL 1_DHS_Group one_F1636819_011.fcs",
  "2016-08-01_PANEL 1_DHS_Group two_F1631374_028.fcs",
  "2016-05-27_PANEL 1_ZF_panel one_F1631054_011.fcs", #test 5 page 4  weird additional population of cells 
  "2016-05-27_PANEL 1_ZF_panel one_F1631061_005.fcs" ,#test 5 page 6  weird additional population of cells
  "2016-05-27_PANEL 1_ZF_panel one_F1631137_009.fcs",#test 5 page 8 lower cluster below the lymphocytes
  "2016-05-27_PANEL 1_ZF_panel one_F1631355_008.fcs" ,#test 5 page 8 lower cluster below the lymphocytes
  "2016-08-01_PANEL 1_DHS_Group two_F1631374_028.fcs",# most over -called
  "2016-08-01_PANEL 1_DHS_Group one_F1636819_011.fcs",# 2nd most over -called
  "2016-07-29_PANEL 1_ZF_Group one_F1632280_001.fcs", # most under -called
  "2016-07-29_PANEL 1_ZF_Group two_F1636833_030.fcs", # 2nd most under -called
  "2016-07-08_PANEL 1_ZF_Group one_F1632339_009.fcs" #Crazy flowrate
  )


# fcsFilesAll = unique(c(fcsFilesAllProbs, fcsFilesAll))

qcOutliers = c(
  "2016-05-13_PANEL 1_ZF_panel one_F1632233_004.fcs",
  "2016-05-20_PANEL 1_ZF_panel one_F1631992_004.fcs",
  "2016-05-20_PANEL 1_ZF_panel one_F1632217_007.fcs",
  "2016-05-26_PANEL 1_ZF_panel one_F1636727_020.fcs",
  "2016-07-08_PANEL 1_ZF_Group one_F1631269_003.fcs",
  "2016-07-08_PANEL 1_ZF_Group one_F1632339_009.fcs",
  "2016-07-08_PANEL 1_ZF_Group one_F1632342_002.fcs",
  "2016-07-27_PANEL 1_ZF_Group one_F1631305_002.fcs",
  "2016-07-27_PANEL 1_ZF_Group one_F1632384_007.fcs",
  "2016-07-29_PANEL 1_ZF_Group one_F1632280_001.fcs",
  "2016-07-29_PANEL 1_ZF_Group one_F1636783_009.fcs",
  "2016-08-01_PANEL 1_DHS_Group one_F1630927_008.fcs",
  "2016-08-01_PANEL 1_DHS_Group two_F1636830_032.fcs"
)
fcsFilesAll = unique(c(fcsFilesAll,fcsFilesAllProbs, qcOutliers))
# fcsFilesAll =qcOutliers

# only use files with gates available
fcsFilesAll = fcsFilesAll[fcsFilesAll %in% mapper$FCS]
# fcsFilesAll = fcsFilesAll[1:5]
fcsFilesAll = split(fcsFilesAll, ceiling(seq_along(fcsFilesAll) / 5))

# mapper[which(mapper$FCS %in% fcsFilesAll),]

# 19:49 -> 20:31
# [30:33]

getStats <- function(gs1, qcVersion, metric, gate) {
  autoStats = getPopStats(gs1, statistic = metric)
  autoStats$METRIC = metric
  autoStats$GATE = gate
  autoStats$QC = qcVersion
  return(autoStats)
}

compFrame <-
  function(frame,
           file,
           gt_lymph,
           d,
           outputDir,
           gateDir,
           qcVersion,
           mapper,
           inputFCSDir) {
    print(paste("compensating ....", file))
    metrics = data.frame()
    comp <- compensation(keyword(frame)$`SPILL`)
    
    chnls <- parameters(comp)
    biexpTrans <-
      flowJo_biexp_trans(
        channelRange = 256,
        maxValue = 262144.0000291775 ,
        pos = 4.418539922,
        neg = 0,
        widthBasis = -100
      )
    tf <- transformerList(chnls, biexpTrans)
    
    print(paste("gating ....", file))
    # gs <- GatingSet(c(new_frame))
    frames = c(frame)
    names(frames) = c(basename(file))
    fs =  as(frames, "flowSet")
    gs1 <- GatingSet(fs)
    gs1 <- compensate(gs1, comp)
    gs1 <- transform(gs1, tf)
    
    gh <- gs1[[1]]
    gating(gt_lymph, gs1)
    
    
    
    
    autoCounts = getStats(gs1 = gs1,
                          qcVersion = qcVersion,
                          metric = "count",
                          "AUTOMATIC")
    # autoProps = getStats(gs1 = gs1,
    #                      qcVersion = qcVersion,
    #                      metric = "freq",
    #                      "AUTOMATIC")
    # metrics = rbind(autoCounts, autoProps)
     metrics = autoCounts

    if (!qcVersion) {
      wsFile = mapper[which(mapper$FCS == file), ]$WSP
      ws <- openWorkspace(wsFile)
      gs <-
        parseWorkspace(
          ws,
          path = inputFCSDir,
          name =1,
          subset = eval(fileToLoad),
          isNcdf = FALSE
        )
      manCounts = getStats(gs1 = gs,
                           qcVersion = qcVersion,
                           metric = "count",
                           "MANUAL")
      metrics = rbind(metrics, manCounts)
      
      # manProps = getStats(gs1 = gs,
      #                     qcVersion = qcVersion,
      #                     metric = "freq",
      #                     "MANUAL")
      # metrics = rbind(metrics, manProps)
      
    }
    print(paste("plotting ....", file))
    result <-
      filter(frame, rectangleGate("remove_from_FR_FM" = c(10000, Inf)))
    overlay.fr2 <- Subset(frame, result)
    t1 = ggcyto(gs1,
                mapping = aes(x = "FSC-A", y = "SSC-A"),
                subset = "root") +
      geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()  + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
    
    # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
    
    t2 = ggcyto(gs1,
                mapping = aes(x = "FSC-A", y = "SSC-A"),
                subset = "boundary") +
      geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()  + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
    
    
    t3 = ggcyto(gs1,
                mapping = aes(x = "FSC-A", y = "SSC-A"),
                subset = "nonDebris") +
      geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()  + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
    # + scale_x_flowJo_biexp()+ scale_y_flowJo_biexp()
    
    
    t4 = ggcyto(gs1,
                mapping = aes(x = "FSC-W", y = "FSC-H"),
                subset = "lymph") +
      geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t5 = ggcyto(gs1,
                mapping = aes(x = "PE-A", y = "FSC-H"),
                subset = "Singlets") +
      geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t6 = ggcyto(gs1,
                mapping = aes(x = "CD3", y = "CD19"),
                subset = "PE-A") +
      geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t1Bcell = ggcyto(gs1,
                     mapping = aes(x = "CD3", y = "CD19"),
                     subset = "CD3-") +
      geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t2Bcell = ggcyto(gs1,
                     mapping = aes(x = "IgD", y = "CD27"),
                     subset = "CD19gate") +
      geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t7 = ggcyto(gs1,
                mapping = aes(x = "CD4", y = "CD8"),
                subset = "CD3+") +
      geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t7_1 = ggcyto(gs1,
                  mapping = aes(x = "CD4", y = "CD8"),
                  subset = "CD4_1") +
      geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t8 = ggcyto(gs1,
                mapping = aes(x = "CD4", y = "HLA-DR"),
                subset = "CD4") +
      geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()
    
    t9 = ggcyto(gs1,
                mapping = aes(x = "CD45RA", y = "CCR7"),
                subset = "CD4") + geom_hex(bins = 100) + ggcyto_par_set(limits = "data") + geom_gate()
    grid.arrange(
      as.ggplot(t1),
      as.ggplot(t2),
      as.ggplot(t3),
      as.ggplot(t4),
      as.ggplot(t5),
      as.ggplot(t6),
      as.ggplot(t1Bcell),
      as.ggplot(t2Bcell),
      as.ggplot(t7),
      as.ggplot(t7_1),
      as.ggplot(t8),
      as.ggplot(t9),
      
      ncol = 2
    )
    # grid.arrange(
    #   as.ggplot(t1)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t2)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t3)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t4)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t5)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t6)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19)
    # )
    #  grid.arrange(
    #   as.ggplot(t1Bcell)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t2Bcell)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t7)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t7_1)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t8)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #   as.ggplot(t9)+ geom_point(data = overlay.fr2, color = "red", cex = 0.5, pch = 19),
    #
    #   ncol = 2
    # )
    # }
    
    # grid.arrange(
    #   as.ggplot(t1),
    #   as.ggplot(t2),
    #   as.ggplot(t3),
    #   as.ggplot(t4),
    #   as.ggplot(t5),
    #   as.ggplot(t6),
    #   as.ggplot(t1Bcell),
    #   as.ggplot(t2Bcell),
    #   as.ggplot(t7),
    #   as.ggplot(t7_1),
    #   as.ggplot(t8),
    #   as.ggplot(t9),
    #
    #   ncol = 2
    # )
    
    d = rbind.data.frame(as.data.frame(getPopStats(gs1)), d)
    outFile <-
      paste(outputDir, gateDir,
            basename(file),
            "_panel1.wsp", sep = "")
    GatingSet2flowJo(gs1, outFile)
    
    QCDEF = FALSE
    if (qcVersion) {
      QCDEF = getIndiceMat(gs1, "QC")
    }
    #     [1] "plotting .... 2016-05-29_PANEL 1_HB_panel one_H_D3_003.fcs"
    # QCDEF
    #  FALSE   TRUE
    # 136918 982260
    print(table(QCDEF))
    gateDef = data.frame(
      QC = QCDEF,
      lymph = getIndiceMat(gs1, "lymph"),
      Singlets = getIndiceMat(gs1, "Singlets"),
      PE_A_Minus = getIndiceMat(gs1, "PE-A"),
      CD3Plus = getIndiceMat(gs1, "CD3+"),
      cd4Pcd8M = getIndiceMat(gs1, "CD4"),
      cd4Mcd8P = getIndiceMat(gs1, "CD8"),
      HLADR = getIndiceMat(gs1, "ActivatedCD4"),
      CCR7PCD45RAP = getIndiceMat(gs1, "CCR7+CD45RA+"),
      CCR7PCD45RAM = getIndiceMat(gs1, "CCR7+CD45RA-"),
      CCR7MCD45RAM = getIndiceMat(gs1, "CCR7-CD45RA-"),
      CCR7MCD45RAP = getIndiceMat(gs1, "CCR7-CD45RA+"),
      BCELL = getIndiceMat(gs1, "CD19gate")
      
    )
    colnames(gateDef) = c(
      "QC_FILTER",
      "lymph",
      "Singlets",
      "PE_A",
      "CD3_PLUS",
      "CD4",
      "CD8",
      "ActivatedCD4",
      "CCR7P_CD45RAP",
      "CCR7P_CD45RAM",
      "CCR7M_CD45RAM",
      "CCR7M_CD45RAP",
      "BCELL"
    )
    # APC-Cy7-A = CD4
    # BUV 395-A =CD8
    
    write.table(
      gateDef,
      sep = "\t",
      quote = FALSE,
      file = paste(
        outputDir,
        gateDir,
        basename(file),
        "_panel1_gate_def.txt",
        sep = ""
      ),
      row.names = FALSE
    )
    return(metrics)
    
  }



# https://www.bioconductor.org/packages/devel/bioc/vignettes/flowCore/inst/doc/HowTo-flowCore.pdf
i = 1
numProcessed = 0
d = data.frame()
metrics = data.frame()
for (files in fcsFilesAll) {
  # fcsFiles = files
  print(files)
  i
  pdf(file = paste(outputDir, "panel1_test5_", i, ".pdf", sep = ""))
  i = i + 1
  for (file in files) {
    fileToLoad <- file
    
    numProcessed = numProcessed + 1
    print(paste("number processed ....", numProcessed))
    print(paste("loading ....", file))
    frame = read.FCS(paste(inputFCSDir, file, sep = ""))
    description(frame)$FILENAME = file
    
    qcFile = paste(tools::file_path_sans_ext(file), ".fcs", sep = "")
    qcFileFull = paste(cleanDir, "QC/", qcFile, sep = "")
    if (!file.exists(qcFileFull)) {
      flow_auto_qc(
        frame,
        folder_results = "QC",
        mini_report = paste(basename(file), "mini"),
        fcs_QC = "",
        pen_valueFS = 50,
        remove_from = "FR_FM"
      )
    }
    
    frame.c = read.FCS(qcFileFull)
    
    metricBase = compFrame(
      frame = frame.c,
      file = file,
      gt_lymph = gt_lymph ,
      d = d,
      outputDir = outputDir,
      gateDir = gateDir,
      qcVersion = FALSE,
      mapper,
      inputFCSDir
    )
    metrics = rbind(metrics, metricBase)
    metricQC = compFrame(
      frame = frame.c,
      file = file,
      gt_lymph = gt_lymphQC ,
      d = d,
      outputDir = outputDir,
      gateDir = gateQCDir,
      qcVersion = TRUE,
      mapper,
      inputFCSDir
    )
    metrics = rbind(metrics, metricQC)
    
  }
  dev.off()
  
}

write.table(
  metrics,
  sep = "\t",
  quote = FALSE,
  file = paste(outputDir, "testTypesTestMetrics5.txt", sep = ""),
  row.names = FALSE
)


write.table(
  d,
  sep = "\t",
  quote = FALSE,
  file = paste(outputDir, "testTypesTest5.txt", sep = ""),
  row.names = FALSE
)
# https://github.com/RGLab/openCyto/issues/110autoplot(gs[[1]])

```


ws <- openWorkspace("/Users/Kitty/temp/fcs/513-520 HRS p1&p2.wsp")
 gs <- parseWorkspace(ws, path = "/Volumes/Beta/data/flow/fcs/", name = 1,subset="2016-05-18_PANEL 1_ZF_panel one_F1632164_009.fcs",isNcdf = FALSE)
t3 = ggcyto(gs,
            mapping = aes(x = "FSC-A", y = "SSC-A"),
            subset = "Lymphocytes (SSC-A v FSC-A)") +
    geom_hex(bins = 200) + ggcyto_par_set(limits = "data") + geom_gate()  + xlim(c(0, 2e5)) + ylim(c(0, 2e5))
write.table(
+     getPopStats(gs),
+     sep = "\t",
+     quote = FALSE,
+     file = "/Volumes/Beta/data/flow/testTypesManual.txt",
+     row.names = FALSE
+ )

```

   
      gt_lymph = rbind(
        gt_lymph,
        add_pop(
        gs1,
        alias = "*",
        pop = "HLADR+/-",
        parent = "CD4",
        dims =  "HLA-DR",
        gating_method = "tailgate",
        gating_args =
        "tol=5e-2",
        preprocessing_method = "standardize_flowset"
        )
        )
      
      gt_lymph = rbind(
        gt_lymph,
        add_pop(
        gs1,
        alias = "*",
        pop = "CD45RA+/-",
        parent = "CD4",
        dims =  "CD45RA",
        gating_method = "mindensity",
        gating_args =
        "gate_range=c(80,150)"
        )
        )
      gt_lymph = rbind(
  gt_lymph,
  add_pop(
    gs1,
    alias = "CCR7_gate",
    pop = "CCR7",
    parent = "CD45RA+",
    dims =  "CCR7",
    gating_method = "flowClust",
    gating_args = "K=2",
    preprocessing_method = "prior_flowClust",
    preprocessing_args = "K=2,min=0"
  )
)


qcDir = "/Volumes/Beta/data/flow/fcs3Clean/QC"
reviewDir = "/Users/Kitty/git/auto-fcs/explore/openCyto/QC/qcReview"
collectResults <-
  list.files(qcDir,
             pattern = "mini.txt$",
             full = TRUE)
df = data.frame()
for (file in collectResults) {
  tmp = read.delim(file,
                   stringsAsFactors = FALSE)
  
  df = rbind(df, tmp)
}

check = df[which((df$X..anomalies.flow.Rate > 20 |
                    df$X..anomalies.Margins > 6) & df$n..of.events > 10000), ]
copy = paste(check$Name.file, ".fcs", sep = "")

